Comparing probabilistic predictive models applied to football

### 2015-07-08 responseToReferee2.tex

\vspace{2mm}
\textbf{Response: we added two paragraphs in subsection 2.2 explaining how the models we are proposing were updated. We used all past results of the teams playing against each other up to the matchday that is being predicted. In this sense, all predictions are ``one-step ahead'' predictions. Regarding the benchmark models, subsection 2.1 comments how they use the information to predict the results. Basically, the predictions of the benchmark models are also one-step ahead predictions, even though they could use the information of the first half of the championship to predict all the second half.} 
% \red{I'm not sure!! Does he mean $1/(\lambda_j\Delta_J^2)$ has to be bounded by the same constant for all $j$'s? Recall that $%\lambda_1>\ldots>\lambda_J$ and $\Delta_1 > \ldots > \Delta_J$}

\vspace{6mm}

\hspace{4mm} \textbf{Reviewer:} \textit{\textbf{Section 1: Introduction.} 
	\begin{itemize}
		\item Line 17 page 2 should ``provisions" be ``predictions"?
		\item It may be beneficial to the reader to briefly describe the second half of the Brazilian
		championship. (As an aside, on page 4 instead of using `` second half", ``second round"
		is used. Are you referring to the same set of games?) How many games per team in the
		second half?
	\end{itemize}}

\vspace{2mm}
\textbf{Response: we changed ``previsions'' to ``predictions''. There was a mistake on p. 4 because we should refer to the ``second round''. We fixed this and added a noted explaining that each team plays 19 matches each round, since the championship has 20 teams. More details on the championship rule are given in Section 4 (Results).}

### 2015-07-08 responseToReferee2.tex

\documentclass[11pt]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath, graphicx, layout, verbatim}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{float}
\usepackage{color}
\usepackage{natbib}
\usepackage{verbatim}
\usepackage{subfig}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[pagebackref,colorlinks=true,urlcolor=blue,citecolor=blue, linkcolor=blue]{hyperref}
\textheight25cm \footskip0.5cm \topmargin-1.5cm \headsep0.2cm
\oddsidemargin-0.5cm \evensidemargin-0.5cm
 \marginparwidth0cm \marginparsep = 0pt
\textwidth16cm

@@

\vspace{2mm}
\textbf{Response:} We thank the referee for taking the time reviewing the paper. Below we list our detailed response to the referee's questions.\\

\vspace{6mm}

\hspace{4mm} \textbf{Reviewer:} \textit{\textbf{General Comments.}
    What the authors have shown in my opinion is that the blackbox predictors they consider as
    competitors do not incorporate covariate information effectively and/or the covariates are not
    informative in learning about game outcomes. Even though it is not the focus of the paper,
    it seems to me that including some detail (at least on an intuitive level) of how information is
    used in these black box methods would be helpful in better understanding why they perform so
    poorly relative to the very unsophisticated MD models.}

\vspace{2mm} \textbf{Response: We included a new subsection
(Subsection 2.1) in Section 2 giving more details about the
benchmark models considered, which are quite similar to many other
approaches in the literature. In fact, we believe that soccer
prediction is really a hard task. This is mainly because several
factors may have a great influence on the final match outcome such
as referee's error, receiving a red card, injury of important
players, weather, among many others. All those factors are difficult
(maybe impossible) to take into account in any statistical model.
Thus, the use of elaborated models leads to small gain in prediction
accuracy when compared to more simpler models, because of the great
unpredictability of soccer matches.}

@@

\vspace{2mm}
\textbf{Response: we added two paragraphs in subsection 2.2 explaining how the models we are proposing were updated. We used all past results of the teams playing against each other up to the matchday that is being predicted. In this sense, all predictions are ``one-step ahead'' predictions. Regarding the benchmark models, subsection 2.1 comments how they use the information to predict the results. Basically, the predictions of the benchmark models are also one-step ahead predictions, even though they could use the information of the first half of the championship to predict all the second half.}
% \red{I'm not sure!! Does he mean $1/(\lambda_j\Delta_J^2)$ has to be bounded by the same constant for all $j$'s? Recall that $%\lambda_1>\ldots>\lambda_J$ and $\Delta_1 > \ldots > \Delta_J$}

\vspace{6mm}

\hspace{4mm} \textbf{Reviewer:} \textit{\textbf{Section 1: Introduction.}
    \begin{itemize}
        \item Line 17 page 2 should ``provisions" be ``predictions"?
        \item It may be beneficial to the reader to briefly describe the second half of the Brazilian
        championship. (As an aside, on page 4 instead of using `` second half", ``second round"
        is used. Are you referring to the same set of games?) How many games per team in the
        second half?
    \end{itemize}}

\vspace{2mm}
\textbf{Response: we changed ``previsions'' by ``predictions''. There was a mistake on p. 4 because we should refer to the ``second round''. We fixed this and added a noted explaining that each team plays 19 matches each round, since the championship has 20 teams. More details on the championship rule are given in Section 4 (Results).}

@@

\vspace{6mm}

\hspace{4mm} \textbf{Reviewer:} \textit{\textbf{Section 2: The models: theoretical background.}
    \begin{itemize}
        \item You point out that the assumptions associated for the MD models and black-box models
        are different. Does this make comparing them untenable? If not, why?
    \end{itemize}}

    \vspace{2mm}
    \textbf{Response: We believe that in the such cases where the model assumptions are quite different, the only possible comparison are in terms
    of predictive accuracy. In the present study, this comparison is possible since all the models report the probabilities of win, draw or loss for the home team.
    Furthermore, from the perspective of soccer fans the most important value of a model is giving ``good'' predictions.}

### 2015-07-08 template.tex
Versao final

	The famous saying of John von Neumann: {\it ``With four parameters I can fit an elephant, and with five I can make him wiggle his trunk''} is a perhaps just a humorous version of Ockham's razor:
	{\it``Entities are not to be multiplied without necessity''}, but illustrates very well a principle that is not often followed by researchers, especially when building predictive models.
	This is true also in the literature of probabilistic sports predictions, particularly regarding models devoted to football or soccer (henceforth football) predictions.

@@

	The predictions of the two benchmark models were published on Internet websites before every matchday, and are treated by us as black-box models (BB1 and BB2). 
	Since one of the authors is responsible for one of these models and the other was described in a dissertation \cite{arruda2000}, we know that both models are based on Holgate bivariate Poisson regression considering as explanatory variables the home field advantage, attack and defense strength for each team, and provide the probabilities of all possible outcomes (home team wins, draws or loses) for each match.
	Because both models are based on elaborate parametric assumptions, they are suitable for the comparison we are proposing: simple versus complex models. To be precise we would have to define what is a ``simple'' model and what is a ``complex'' model. Based on the opening quotations, one may say that the number of parameters defines the complexity of a model, which is not exactly the definition we will use, since we consider other criteria, like the interpretation of explanatory variables and mathematical structure of the statistical model.

@@

	Formally, let $X$ be a random variable
	taking values in $\mathcal{X}=\{1,2,3\}$ indicating
	the outcome of the match, with 1 standing for home win, 2 for draw and 3 for away win.  Moreover, let $P=(P_1,P_2,P_3)$ denote one's probabilistic prediction
	about $X$, i.e., $P$ lies in the 2-simplex set $\Delta_2=\{(p_1,p_2,p_3):p_1+p_2+p_3=1, \ p_1,p_2,p_3\geq0\}$ (see Figure
	\ref{fig:simplex}). 	
	A scoring rule is a function 
	that assigns a real number (score) $S(x,P)$ to each $x \in \mathcal{X}$
	and $P \in \Delta_2$ 
	such that 
	for any given $x$ in $\mathcal{X}$, the score  $S(x,P)$ is minimized when $P$ is 
	$(1,0,0)$, $(0,1,0)$ or $(0,0,1)$ depending if $x$ is 1, 2 or 3, respectively.
	The score $S(x,P)$ can be thought as
	a penalty to be paid when one assigns the 
	probabilistic prediction $P$ and outcome
	$x$ occurs. Also, the ``best'' possible score (i.e., the smallest score value) is achieved when the probabilistic prediction for the outcome of the game is perfect. A scoring rule can also be defined to be such that a \emph{large} value of the score indicates good forecasts.
	%Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S(x,P_X)=0$ for each $x$.

@@

	Say a model's probabilistic prediction for $X$ is $P^*=(P_1^*,P_2^*,P_3^*).$ A proper scoring rule $S$ is a scoring rule  such that the mean score value 
	$$E_{P^*}[S(X,P)]=\sum_{x=1}^3 S(x,P)P^*_x$$
	 is minimized when $P=P^*$.
	In other words, if one announces
	$P$ as his probabilistic prediction
	and $S$ as a scoring rule, the lowest
	expected penalty is obtained by reporting the correct model's  uncertainty about $X$, $P^*$.
	Thus, the use of a proper scoring rule encourages the forecaster to announce $P^*$ (the correct one)
	as his probabilistic prediction  rather than some other quantity. 
	%Moreover, this method automatically leads to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events and, therefore, the accumulated score may be considered a concrete measure of success of each model.

@@

	Although our benchmark models were here defined as complex, we can find much more elaborate models in the literature, hindering the estimation and prediction processes.
	Among them, we can cite those based on modeling the match as a stochastic process evolving through time \citep{Dixon98, Volf2009, Titman2015}, those allowing for the team performance parameters to change along the season \citep{Rue2000, Crowder2002, Owen2011, Koopman2015} and those modeling dependence between number of goals by means of bivariate count distributions \citep{Dixon97, Karlis2003, McHale2007, McHale2011}.
	However, models build to predict the number of goals must predict a more specific outcome and this usually requires more elaborate modeling.

	According to standard scoring rules and other criteria here adopted to qualify the predictions, our findings show that the predictive power of the three simple models was quite similar to the benchmark models.
	This work thus indicates that while there has been a boom in the literature with the proposal of several fancy and involved models, especially in the era of big data, simple models can many times do as well as their complex counterparts.

	This is often unnoticed by users of such methods, which leads to the overuse of models that are
	hard to interpret and to implement, and that also require one to keep track of several covariates. We would call this phenomenon the {\it king's nudity}, in allusion to Anderson's {\it The Emperor's New Clothes}, \cite{emperor}: ``But he isn't wearing anything at all!''. However , we emphasize that this phenomenon is not always observed. 
	In this spirit, since we have explored only the prediction of football matches, we would like to study if the same extends to other fields such as, for instance, weather or economics forecasts.

>>>>

    The famous saying of John von Neumann: {\it ``With four parameters I can fit an elephant, and with five I can make him wiggle his trunk''} is a perhaps just a humorous version of Ockham's razor:
    {\it``Entities are not to be multiplied without necessity''}, but illustrates very well a principle that is not often followed by researchers, especially when building predictive models.
    This is true also in the literature of probabilistic sports predictions, particularly regarding models devoted to football or soccer (henceforth, football) predictions.

@@

    The predictions of the two benchmark models were published on Internet websites before every matchday, and are treated by us as black-box models (BB1 and BB2).
    Since one of the authors is responsible for one of these models and the other was described in a dissertation \citep{arruda2000}, we know that both models are based on Holgate bivariate Poisson regression considering as explanatory variables the home field advantage, attack and defense strength for each team, and provide the probabilities of all possible outcomes (home team wins, draws or loses) for each match.
    Because both models are based on elaborate parametric assumptions, they are suitable for the comparison we are proposing: simple versus complex models. To be precise, we would have to define what is a ``simple'' model and what is a ``complex'' model. Based on the opening quotations, one may say that the number of parameters defines the complexity of a model, which is not exactly the definition we will use, since we consider other criteria, like the interpretation of explanatory variables and mathematical structure of the statistical model.

@@

    Formally, let $X$ be a random variable
    taking values in $\mathcal{X}=\{1,2,3\}$ indicating
    the outcome of the match, with 1 standing for home win, 2 for draw and 3 for away win.  Moreover, let $P=(P_1,P_2,P_3)$ denote one's probabilistic prediction
    about $X$, i.e., $P$ lies in the 2-simplex set $\Delta_2=\{(p_1,p_2,p_3):p_1+p_2+p_3=1, \ p_1,p_2,p_3\geq0\}$ (see Figure
    \ref{fig:simplex}).
    A scoring rule is a function
    that assigns a real number (score) $S(x,P)$ to each $x \in \mathcal{X}$
    and $P \in \Delta_2$
    such that
    for any given $x$ in $\mathcal{X}$, the score  $S(x,P)$ is minimized when $P$ is
    $(1,0,0)$, $(0,1,0)$ or $(0,0,1)$ depending if $x$ is 1, 2 or 3, respectively.
    The score $S(x,P)$ can be thought as
    a penalty to be paid when one assigns the
    probabilistic prediction $P$ and outcome
    $x$ occurs. Also, the ``best'' possible score (i.e., the smallest score value) is achieved when the probabilistic prediction for the outcome of the game is perfect. A scoring rule may also be defined to be such that a large value of the score indicates good forecasts.
        %Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S(x,P_X)=0$ for each $x$.

@@

    Say a model's probabilistic prediction for $X$ is $P^*=(P_1^*,P_2^*,P_3^*).$ A proper scoring rule $S$ is a scoring rule  such that the mean score value
    $$E_{P^*}[S(X,P)]=\sum_{x=1}^3 S(x,P)P^*_x$$
     is minimized when $P=P^*$.
    In other words, if one announces
    $P$ as his probabilistic prediction
    and uses $S$ as scoring rule, the lowest
    expected penalty is obtained by reporting the correct model's  uncertainty about $X$, $P^*$.
    Thus, the use of a proper scoring rule encourages the forecaster to announce $P^*$ (the correct one)
    as his probabilistic prediction  rather than some other quantity.
    %Moreover, this method automatically leads to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events and, therefore, the accumulated score may be considered a concrete measure of success of each model.

@@

    Although our benchmark models were here defined as complex, we can find much more elaborate models in the literature, hindering the estimation and prediction processes.
    Among them, we can cite those based on modeling the match as a stochastic process evolving through time \citep{Dixon98, Volf2009, Titman2015}, those allowing for the team performance parameters to change along the season \citep{Rue2000, Crowder2002, Owen2011, Koopman2015} and those modeling dependence between number of goals by means of bivariate count distributions \citep{Dixon97, Karlis2003, McHale2007, McHale2011}. Models that predict the number of goals have more specific objectives than those prediciting the final outcomes and therefore require more elaborate modeling, justifying the complexity of such models.

    According to standard scoring rules and other criteria here adopted to qualify the predictions, our findings show that the predictive power of the three simple models was quite similar to the benchmark models.
    %We compared the probabilistic predictions of three simple models with the ones provided by two more complex models for the second round of the 2014 Brazilian football championships of first and second divisions.
    This work thus indicates that while there has been a boom in the literature with the proposal of several fancy and involved models, especially in the era of big data, simple models can many times do
    as well as their complex counterparts.

    This is often unnoticed by users of such methods, which leads to the overuse of models that are
    hard to interpret and to implement, and that also require one to keep track of several covariates. We would call this phenomenon the {\it king's nudity} in allusion to Anderson's {\it The Emperor's New Clothes}, \cite{emperor}: ``But he isn't wearing anything at all!''. However, this is not always the case. In this spirit, since we have explored only the prediction of football matches, we would like to study if the same extends to other fields such as, for instance, weather or economics forecasts.
